Data Analysis

### 2023-03-29 ms1_data_analysis.Rmd
Before introducing the phylogenetic correlation to the deltaBiomass models.

```{r}
# Model explorations
simple.sla <- gls(
  lrrtb ~ lsla.d*site + I(lsla.d^2)*site + richness + labu + lrrsla,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

@@

# Model selection.
# 1. Full model (quadratic, and site interactions) vs no traits
anova(no.traits, full.model) # full model is much better
anova(full.model, full.model.noint)

@@

# WATER
# Check quadratic
anova(full.model.noint, just.water.quadratic.noint) # Model with just water content

# Is the quadratic form necessary
anova(just.water.quadratic.noint, just.water.linear.noint) # no difference but quadratic has lower AIC

@@

>>>>

```{r, results='asis'}
# Model explorations
simple.sla <- gls(
  lrrtb ~ lsla.d*site + I(lsla.d^2)*site + richness + labu + lrrsla,
                data = deltaData %>% filter(unified_names %in% myTree.rand.poly$tip.label),
  method = "ML"
  )

@@

# Model selection.
# 1. Full model (quadratic, and site interactions) vs no traits
# anova(no.traits, full.model) # full model is much better
# anova(full.model, full.model.noint)

@@

# WATER
# Check quadratic
# anova(full.model.noint, just.water.quadratic.noint) # Model with just water content

# Is the quadratic form necessary
# anova(just.water.quadratic.noint, just.water.linear.noint) # no difference but quadratic has lower AIC

@@

# Predators

```{r}
subsetVars <- c("c", "p")

# Model selection.
# 1. Full model (quadratic, and site interactions) vs no traits
# anova(no.traits, full.model)            # full model is better
# anova(full.model, full.model.noint)     # No difference, so we can drop the interaction

# 2. Is the quadratic relationship significant
# anova(full.model.noint, all.linear)            # Quadratic is a bit better
# anova(full.model.noint, all.linear.noint)      # All linear no int is more efficient

# 3. REMOVE SLA

# anova(full.model.noint, q.herb.q.water.noint) # withouth sla is slightly better (AIC)
# anova(full.model.noint, q.herb.q.water)
# anova(full.model.noint, q.herb.l.water.noint) # Linear water and quadratic herbivory works even better
# anova(full.model.noint, q.herb.l.water) # interactions make it worse
# anova(full.model.noint, l.herb.l.water.noint) # Full model still better
# anova(q.herb.l.water.noint, l.herb.l.water.noint) # All linear? Slightly better
# anova(full.model.noint, l.herb.l.water)

# All linear with interaction seems to work best, but the interaction is not improving the model so I will keep it out.

# Including SLA did not mprove the model significantly
# tab_model(l.herb.l.water.noint)
# anova(no.traits, l.herb.l.water.noint)

# But if we were to start first by remving perc.herbivory?
# anova(full.model.noint, q.sla.q.water.noint) # Model without herbivory is not better
# anova(full.model.noint, q.sla.q.water)       # Interaction makes it worse

# anova(full.model.noint, q.sla.l.water.noint)   # Full model is still better
# anova(full.model.noint, q.sla.l.water)         # Full model wins again
# anova(full.model.noint, l.sla.l.water.noint)   # And again by AIC criterium
# anova(full.model.noint, l.sla.l.water)         # And again

# The above suggests that models without hebivory and sla do better job. 
# Water content is again the key trait that helps predict the change in biomass in response to the treatment.

# WATER
# Check quadratic
# anova(l.herb.l.water.noint, just.water.quadratic.noint) 
# Model with just water content is worse than that without herbivores

# Is the quadratic form necessary
# anova(l.herb.l.water.noint, just.water.linear.noint) # Herbivory is an important trait 
# anova(l.herb.l.water.noint, just.herb.linear.noint) # Water is an important trait, both should stay in the model

# So the best model is the one without the traits!
tab_model(l.herb.l.water.noint)
